Optimal updating magnitude in adaptive flat-distribution-sampling simulations

### 2017-08-25 doc/invt.tex
add scheme.pdf (in doc/fig/figs.tex)
invt.tex: V8.1

\documentclass[reprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
%\usepackage{tikz}
%\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

@@

\subsubsection{\label{sec:Gaussian_scheme}
Gaussian and homogeneous updating schemes}

@@

@@

Now that we can compute the optimal schedule
for a given updating scheme,
we turn to the problem of finding
the optimal updating scheme.

@@

We will consider two special classes of bandpass updating schemes,
homogeneous and polynomial bandpass updating schemes.
%
First,
by applying this construction to the basis of cosine and sine functions,
we get a homogeneous bandpass updating scheme
that resembles the Gaussian updating scheme,
but with long-range wiggles
[cf. Appendix \ref{sec:homo_bandpass}
and Fig. \ref{fig:mat}(b)
for an example].
%
In Sec. \ref{sec:lj},
we will further compare the performance of
the homogeneous bandpass and Gaussian updating schemes
in an example.
%
%While both updating scheme are able to deliver
%a smooth and differentiable potential of mean force,
%the bandpass updating scheme is superior in asymptotic convergence,
%because of the set of clear-cut spectrum of eigenvalues.
%Eq. \eqref{eq:lambda_bandpass},
%instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
%
%\footnote{As an estimate of error,
%  we note that truncating a linear or quadratic function
%  at mode $K$ for a non-periodic variable
%  results in an error roughly proportional to
%  $|\max_i u_i - \min_i u_i|^2/K^3$.}
%
The polynomial bandpass updating schemes will be discussed below.

@@

The optimal schedule was more effective
than the inverse-time schedule
in reducing the error,
as shown in Table \ref{tab:lj_error}.
%
For this case,
we further decomposed the error
into the eigenmodes
by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl},
and plotted the error components, $\langle \tilde v_k^2 \rangle$,
in Fig. \ref{fig:lj_xerr}.
%
We found that the inverse-time schedule tended to
reduce the updating magnitude too quickly,
leaving a hump on the error profile
for the $k = 10$ to $25$ modes.
%
By a delayed reduction,
the optimal schedule more effectively reduced
the error in these modes,
and produced a flatter profile for the final error
(the remaining unevenness of the profile
was likely due to the error of estimated values of $\Gamma_k$'s).

@@

For the Gaussian updating scheme,
the error depended not only the schedule
but also the width of the Gaussian.
%
Although widening the Gaussian
can suppress local noise,
it can also mask finer features of the PMF.
%it also diminishes the eigenvalues
%[by Eq. \eqref{eq:lambda_Gaussian}]
%of more middle-range modes
%(which often represent finer features of the PMF)
%and hence slows down the error reduction there
%[by Eq. \eqref{eq:vt_diffeq_mbin}].
%
Thus, there is an optimal width that balances the two factors.
%
According to the analytical prediction
(assuming the optimal schedule),
the optimal width was around $\sigma = 0.29$ in this case,
and it would decrease with the simulation length
and approach the limit of single-bin updating scheme,
$\sigma \to 0$, asymptotically.

@@

As another proof of principle,
we tested the adaptive Gaussian ensemble method
introduced in Sec. \ref{sec:age}
on the two-dimensional $L\times L$ ten-state
Potts model\cite{wu1982, newman, wang2001, wang2001pre}
in the energy space.
%
In this case, % $\Phi(\mathbf x) = 0$ and
the intrinsic distribution is
the density of states, $g(E)$.
%
We spread the Gaussian umbrellas
of the same width $\sigma_E = L$ over $[-1.8L^2, -0.8L^2]$
with an even spacing of $L$.
%
In attempting inter-umbrella transitions,
the target umbrella was chosen from
any of the other umbrellas.
%
Initially, we set
$c_1^{(i)} = \tilde \beta_c \, \sigma_E$
for all umbrellas,
with $\tilde \beta_c = 1.4$
being the approximate inverse critical temperature.
%
For configuration sampling in each umbrella,
we used the Wolff algorithm\cite{wolff1989, newman}
(which is more efficient than the Metropolis algorithm)
at the inverse temperature, $\beta = c_1/\sigma_E$,
followed by a Metropolis step with
acceptance probability
$\min\bigl\{1, e^{c_2 [R_2(E) - R_2(E')]} \bigr\}$,
where $E$ and $E'$ are the energies of the old and new states.
%
The acceptance ratio was no less than 92\% in any umbrella.
%
In the initial WL stages,
the initial updating magnitude was $0.01$;
the fluctuation threshold for stage transitions
was $F = 0.2$ [using the definition given by Eq. \eqref{eq:F2st}];
and the reduction factor was $1/2$.

@@

Below we use the transition matrix method
to estimate $\Gamma_k$ in a model for local diffusion\cite{bussi2006}.
%
We assume that the propagation along the collective-variable is Markovian
such that the evolution can be approximately
described by a transition matrix, $\mathbf A$, with
\begin{align*}
\langle h_i(t) \rangle = \rho_i,
\quad
\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
\quad \mbox{(for $t \ge 0$)}
.
\end{align*}
Further, if the transition matrix % satisfies detailed balance
has a full set of orthonormal eigenvectors, $\psi_{li}$
%that are orthonormal with respect to $\pmb\rho$
[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
\eqref{eq:eig_orthonormal_rows}],
it can be decomposed as
$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
with $\Lambda_l$ being the eigenvalue of mode $l$.
%
The autocorrelation function can then be computed as
%
\begin{align*}
\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
=
\sum_{l = 0}^{n-1} \Lambda_l^t \,
\left(
  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
\right)^2
.
\end{align*}
%
In particular, if the eigenvectors, $\pmb \phi_k$
and $\pmb \psi_k$, coincide,
we would have
$\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
=
\Lambda_k^t,$
and
\begin{equation}
  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  = \frac{1+\Lambda_k}{1-\Lambda_k}
  .
  \notag
  %\label{eq:Gamma_est}
\end{equation}
%
To model a local sampling process,
we can borrow Eq. \eqref{eq:lambda_Gaussian} for $\Lambda_k$,
with $\sigma_z$ reinterpreted as
the average move size along $z$ during sampling,
then
\begin{align}
\Gamma_k = \coth\left[
  \left(
    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
  \right)^2
  \right].
  \notag
  %\label{eq:Gamma_local}
\end{align}
Thus, we have $\Gamma_k \propto \Delta z^2/k^2$
for small $k$,\cite{bussi2006}
whereas $\Gamma_k \to 1$ as $k$ increases.
%
However, this method tends to underestimate
the first few $\Gamma_k$'s.
%which means that doubling the sampling range, $\Delta z$,
%would make the histogram flattening
%at least four times as difficult,
%i.e. FDS methods are most effective
%for short ranges\cite{wang2001, wang2001pre}.
%
%Using finer bins would, however, not change this value.
%

@@

For a periodic variable\cite{dama2014},
the element of the updating matrix, $w_{ij}$,
can only be a function of the difference of indices, $i-j$, modulo $n$,
and we write it as
%
\begin{equation}
  w_{ij}
  =
  \mu_{i-j}
  +
  \mu_{i-j+n}
  +
  \mu_{i-j-n}
  ,
  \notag
  %\label{eq:w_band_pbc}
\end{equation}
%
for some numbers, $\mu_l$'s, which
define an unadjusted window function.
%
For simplicity, we will assume that
$\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
and they can be cutoff at $l = b$,
where $b$ is the integral part of $(n-1)/2$,
i.e. $\mu_l = 0$ for $l > b$.
%
\note{The cutoff $b$ is a bit tricky to define
in order to encompass the cases of even and odd $n$.}%
%
To satisfy Eq. \eqref{eq:w_sumj}, we further impose
%
\begin{equation}
  \mu_{-b} + \cdots + \mu_b = 1
  .
  \notag
  %\label{eq:mu_normalization}
\end{equation}

@@

The Gaussian updating scheme
is defined by a Gaussian-shaped $\mu_l$.
%
This updating scheme is commonly
used in metadynamics.
%
Below we show its convergence
in the continuous limit,
and how to maintain the convergence
on a finite grid.

@@

If the width of the Gaussian, $\sigma_z$,
is much less than $\Delta z$,
then we can extend the limits of the integrals
to infinity, and normalize $\mu(z)$ as
%
\begin{equation}
  \mu(z)
  \approx
  \frac{            1            }
       { \sqrt{ 2 \pi } \sigma_z }
  %
  \exp\left(
        -\frac{   z^2   }
              { 2 \, \sigma_z^2 }
      \right)
  .
\notag
\end{equation}
%
Then, from Eq. \eqref{eq:wband_eigenvalue},
we get Eq. \eqref{eq:lambda_Gaussian},
which shows that all eigenvalues are positive.
%
However,
when discretized on a finite grid,
some eigenvalues can turn negative.
%
To maintain the positivity of eigenvalues,
we can redefine the Gaussian scheme
from the same spectrum of eigenvalues, Eq. \eqref{eq:lambda_Gaussian},
and then inversely computing the window function
using Eq. \eqref{eq:mu_from_lambda},
or the updating matrix using Eq. \eqref{eq:w_from_phi}.
%
An example is shown in Fig. \ref{fig:mat}(a).

@@

For a non-periodic variable,
we have $\lambda_n = (-1)^{n-K}$
from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan},
and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \mathrm{sinr}(K, l, 2 \, n)$.
%\begin{equation}
%  \mu_l
%  =
%  (-1)^{n-K+l} / ( 2 \, n )
%    +
%  \mathrm{sinr}(K, l, 2n)
%  .
%  \notag
%  %\label{eq:mu_sinc_refl}
%\end{equation}
%
The updating matrix is given by
$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.
%
\note{$w_{ij}$
  is free from the oscillatory term, $(-1)^{n-K+1}$,
  that appeared in the window function, since
  its contribution to $\mu_{i-j}$ to the updating matrix
  is canceled by its contribution to either $\mu_{i+j-1}$
  or $\mu_{i+j-2n-1}$,
  whichever exists, of the opposite sign.}
%
The homogeneous bandpass updating scheme
resembles the Gaussian updating scheme
except for some long-range wiggles,
as shown in Fig. \ref{fig:mat}(b).
%

@@

\note{The table of symbols is listed in Table \ref{tab:symbols}.
  \begin{table*}
  \footnotesize
  \centering
  \rowcolors{1}{white}{LightGray}
  \setlength{\tabcolsep}{4pt} % column separation
  \caption{\label{tab:symbols}
    Table of symbols.}
  \begin{tabular}{l | p{12cm} }
    Symbol          &   Description \\
    \hline
    $\mathbf{A}$    &   Transition matrix. \\
    $\alpha$        &   Updating magnitude. \\
    $b$             &   Cutoff of the window function of homogeneous updating schemes. \\
    $C_M$           &   Integral of the mass function, $M(\bar q)$.  \\
    $\delta$        &   Dirac's or discrete $\delta$-function. \\
    $c$             &   Constant. \\
    $\Err, \Err_R, \Err_A$          &   Error, residual error, asymptotic error. \\
    $\ln f$         &   Updating factor in the WL algorithm.  \\
    $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
    $g$             &   $1$ for a periodic variable, or for $2$ a non-periodic one. \\
    $G_i$           &   Integral of the autocorrelation function of histogram fluctuation at bin $i$. \\
    $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
    $h_i(t)$        &   Instantaneous histogram.  \\
    $H_i$           &   Average normalized histogram.  \\
    $i, j$          &   Bin indices. \\
    $\lambda_k$     &   The $k$th eigenvalue of the updating matrix. \\
    $k, l$          &   Mode indices. \\
    $K$             &   Cutoff wave number of bandpass updating schemes.  \\
    $M(q)$          &   Mass function.   \\
    $m(q)$          &   Mass distribution,
                        $m(q) \equiv M(q)/\int_0^{ q(T) } M(q') \, dq'$.  \\
    $\mu_i$         &   Updating window function of a homogeneous updating scheme. \\
    $n$             &   Number of bins. \\
    $\nu_k$         &   $\nu_k \equiv \lambda_k / \lambda$. \\
    $\rho_i$        &   Flat sampling distribution. \\
    $\pi_i$         &   (Instantaneous) distribution. \\
    $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
    $\bar q(t)$     &   $q(T) - q(t)$.  \\
    $s, S$          &   Stage index and the total number of stages in the Wang-Landau algorithm. \\
    $\sigma_{ij}(t)$   &   Autocorrelation function of $\xi_i(t)$. \\
    $t, \tau$       &   Time. \\
    $T$             &   Simulation length. \\
    $\vartheta_k(q')$       &   $\vartheta_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
    $u_i(t)$        &   Original bias potential. \\
    $v_i(t)$        &   Shifted bias potential. \\
    $\mathbf w$     &   Updating matrix. \\
    $v_{*i}(t)$     &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
    $\xi_{*i}(t)$ &   The noise of histogram. \\
    ${\tilde v}_k$  &   Mode of the shifted bias potential. \\
    ${\tilde u}_k$  &   Mode of the original bias potential. \\
    $z$             &   Reaction coordinate. \\
    $\xi_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
  \end{tabular}
  \end{table*}
}

>>>>

\documentclass[preprint, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[table,usenames,dvipsnames]{xcolor}
%\usepackage{tikz}
%\usetikzlibrary{calc}
\usepackage{hyperref}
\usepackage{setspace}

@@

\subsubsection{\label{sec:Gaussian_scheme}
Gaussian updating scheme}

@@

\subsubsection{\label{sec:homo_scheme}
Homogeneous updating schemes}

@@

Now that we can compute the optimal schedule
for a given updating scheme,
we turn to the problem of optimizing
the updating scheme.

@@

We will consider two special classes of bandpass updating schemes,
homogeneous and polynomial bandpass updating schemes.
%
A summary of updating schemes is given in Fig. \ref{fig:scheme}.
%
First,
by applying this construction to the basis of cosine and sine functions,
we get a homogeneous bandpass updating scheme
that resembles the Gaussian updating scheme,
but with long-range wiggles,
as shown in Fig. \ref{fig:mat}(b)
(cf. Appendix \ref{sec:homo_bandpass}).
%
In Sec. \ref{sec:lj},
we will further compare the performance of
the homogeneous bandpass and Gaussian updating schemes
in an example.
%
%While both updating scheme are able to deliver
%a smooth and differentiable potential of mean force,
%the bandpass updating scheme is superior in asymptotic convergence,
%because of the set of clear-cut spectrum of eigenvalues.
%Eq. \eqref{eq:lambda_bandpass},
%instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
%
%\footnote{As an estimate of error,
%  we note that truncating a linear or quadratic function
%  at mode $K$ for a non-periodic variable
%  results in an error roughly proportional to
%  $|\max_i u_i - \min_i u_i|^2/K^3$.}
%
The polynomial bandpass updating schemes will be discussed below.

\begin{figure}[h]\centering
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/scheme.pdf}
  }
  \caption{
    \label{fig:scheme}
    Summary of updating schemes.
    %
    Only the Gaussian updating scheme
    require explicit calculation for the optimal schedule,
    because the optimal schedule of bandpass updating schemes
    is always given by inverse-time formula.
    %
    The target distributions of all updating schemes are flat except
    that for the adaptive Gaussian ensemble method,
    which is a Gaussian distribution
    or a mixture of Gaussian distributions.
    \note{Do we need these descriptions, or the figure at all?}%
  }
\end{figure}

@@

The optimal schedule was more effective
than the inverse-time schedule
in reducing the error,
as shown in Table \ref{tab:lj_error}.
%
We further decomposed the error
into the eigenmodes
by Eqs. \eqref{eq:gft_def} and \eqref{eq:wband_eigenvector_refl},
and plotted the error components, $\langle \tilde v_k^2 \rangle$,
in Fig. \ref{fig:lj_xerr}.
%
We found that the inverse-time schedule tended to
reduce the updating magnitude too quickly,
leaving a hump on the error profile
for the $k = 10$ to $25$ modes.
%
By a delayed reduction,
the optimal schedule more effectively reduced
the error in these modes,
and produced a flatter profile for the final error
(the remaining unevenness of the profile
was likely due to the error of estimated values of $\Gamma_k$'s).

@@

For the Gaussian updating scheme,
the error depended not only the schedule
but also the width of the Gaussian.
%
Although widening the Gaussian
can suppress local noise,
it can also mask finer features of the PMF.
%it also diminishes the eigenvalues
%[by Eq. \eqref{eq:lambda_Gaussian}]
%of more middle-range modes
%(which often represent finer features of the PMF)
%and hence slows down the error reduction there
%[by Eq. \eqref{eq:vt_diffeq_mbin}].
%
Thus, there is an optimal width that balances the two factors.
%
According to the analytical prediction,
the optimal width (assuming the optimal schedule)
was around $\sigma = 0.29$ in this case,
and it would decrease with the simulation length
and approach the limit of single-bin updating scheme,
$\sigma \to 0$, asymptotically.

@@

As another proof of principle,
we tested the adaptive Gaussian ensemble method
introduced in Sec. \ref{sec:age}
on the two-dimensional $L\times L$ ten-state
Potts model\cite{wu1982, newman, wang2001, wang2001pre}
in the energy space.
%
In this case, % $\Phi(\mathbf x) = 0$ and
the intrinsic distribution is
the density of states, $g(E)$.
%
We spread the Gaussian umbrellas
of the same width $\sigma_E = L$ over $[-1.8L^2, -0.8L^2]$
with an even spacing of $L$.
%
In attempting inter-umbrella transitions,
the target umbrella is randomly chosen from
either one of the nearest neighbors ($50\%$ chance),
or any other umbrella.
%
\note{Here, we need to rerun simulations.}%
%
Initially, we set
$c_1^{(i)} = \tilde \beta_c \, \sigma_E$
for all umbrellas,
with $\tilde \beta_c = 1.4$
being the approximate inverse critical temperature.
%
For configuration sampling in each umbrella,
we used the Wolff algorithm\cite{wolff1989, newman}
(which is more efficient than the Metropolis algorithm)
at the inverse temperature, $\beta = c_1/\sigma_E$,
followed by a Metropolis step with
acceptance probability
$\min\bigl\{1, e^{c_2 [R_2(E) - R_2(E')]} \bigr\}$,
where $E$ and $E'$ are the energies of the old and new states.
%
The acceptance ratio was no less than 92\% in any umbrella.
%
In the initial WL stages,
the initial updating magnitude was $0.01$;
the fluctuation threshold for stage transitions
was $F = 0.2$ [using the definition given by Eq. \eqref{eq:F2st}];
and the reduction factor was $1/2$.

@@

Below we use the transition matrix method
to estimate $\Gamma_k$ in a model for local diffusion\cite{bussi2006}.
%
We assume that the propagation along the collective-variable is Markovian
such that it can be approximately
described by a transition matrix, $\mathbf A$, with
\begin{align*}
\langle h_i(t) \rangle = \rho_i,
\quad
\langle h_i(t) \, h_j(0) \rangle = \bigl(\mathbf A^t\bigr)_{ij} \, \rho_j
\quad \mbox{(for $t \ge 0$)}
.
\end{align*}
Further, if the transition matrix % satisfies detailed balance
has a complete set of orthonormal eigenvectors, $\psi_{li}$
%that are orthonormal with respect to $\pmb\rho$
[cf. Eqs. \eqref{eq:eig_orthonormal_cols} and
\eqref{eq:eig_orthonormal_rows}],
it can be decomposed as
$A_{ij} = \sum_{l=0}^{n-1} \Lambda_l \, \psi_{li} \, \psi_{lj} / \rho_j$,
with $\Lambda_l$ being the eigenvalue of mode $l$.
%
The autocorrelation function can then be computed as
%
\begin{align*}
\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
=
\sum_{l = 0}^{n-1} \Lambda_l^t \,
\left(
  \sum_{i=1}^n \frac{ \phi_{ki} \, \psi_{li} }{ \rho_i }
\right)^2
.
\end{align*}
%
In particular, if the eigenvectors, $\pmb \phi_k$
and $\pmb \psi_k$, coincide,
we would have
$\bigl\langle
  \tilde f_k(t) \, \tilde f_k(0)
\bigr\rangle
=
\Lambda_k^t,$
and
\begin{equation}
  \Gamma_k = 1 + 2 \, \sum_{t=1}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  = \frac{1+\Lambda_k}{1-\Lambda_k}
  .
  \notag
  %\label{eq:Gamma_est}
\end{equation}
%
To model a local sampling process,
we can borrow Eq. \eqref{eq:lambda_Gaussian} for $\Lambda_k$,
with $\sigma_z$ reinterpreted as
the average move size along $z$ during sampling,
then
\begin{align}
\Gamma_k = \coth\left[
  \left(
    \frac{ \pi \, k \, \sigma_z^\mathrm{(samp)} } { g \, \Delta z}
  \right)^2
  \right].
  \notag
  %\label{eq:Gamma_local}
\end{align}
Thus, we have $\Gamma_k \propto \Delta z^2/k^2$
for small $k$,\cite{bussi2006}
whereas $\Gamma_k \to 1$ as $k$ increases.
%
However, this method tends to underestimate
the first few $\Gamma_k$'s.
%which means that doubling the sampling range, $\Delta z$,
%would make the histogram flattening
%at least four times as difficult,
%i.e. FDS methods are most effective
%for short ranges\cite{wang2001, wang2001pre}.
%
%Using finer bins would, however, not change this value.
%

@@

For a periodic variable\cite{dama2014},
the element of the updating matrix, $w_{ij}$,
can only be a function of the difference of indices, $i-j$, modulo $n$,
and we write it as
%
\begin{equation}
  w_{ij}
  =
  \mu_{i-j}
  +
  \mu_{i-j+n}
  +
  \mu_{i-j-n}
  ,
  \notag
  %\label{eq:w_band_pbc}
\end{equation}
%
for some numbers, $\mu_l$'s, which
define an unadjusted window function.
%
For simplicity, we will assume that
the $\mu_l$'s are symmetric, $\mu_l = \mu_{-l}$,
and they can be cutoff at $l = b$,
where $b$ is the integral part of $(n-1)/2$,
i.e. $\mu_l = 0$ for $l > b$.
%
\note{The cutoff $b$ is a bit tricky to define
in order to encompass the cases of even and odd $n$.}%
%
To satisfy Eq. \eqref{eq:w_sumj}, we further impose
%
\begin{equation}
  \mu_{-b} + \cdots + \mu_b = 1
  .
  \notag
  %\label{eq:mu_normalization}
\end{equation}

@@

The Gaussian updating scheme
is defined by a Gaussian-shaped window function, $\mu_l$.
%
This updating scheme is commonly
used in metadynamics.
%
Below we show its convergence
in the continuous limit,
and how to maintain the convergence
on a finite grid.

@@

If the width of the Gaussian, $\sigma_z$,
is much less than $\Delta z$,
then we can extend the limits of the integrals
to infinity, and normalize $\mu(z)$ as
%
\begin{equation}
  \mu(z)
  \approx
  \frac{            1            }
       { \sqrt{ 2 \pi } \sigma_z }
  %
  \exp\left(
        -\frac{   z^2   }
              { 2 \, \sigma_z^2 }
      \right)
  .
\notag
\end{equation}
%
Then, from Eq. \eqref{eq:wband_eigenvalue},
we get Eq. \eqref{eq:lambda_Gaussian},
which shows that all eigenvalues are positive.
%
However,
when discretized on a finite grid,
some eigenvalues can turn negative.
%
To maintain the positivity of eigenvalues,
we can redefine the Gaussian scheme
from the same spectrum of eigenvalues, Eq. \eqref{eq:lambda_Gaussian},
and then inversely computing the window function
using Eq. \eqref{eq:mu_from_lambda},
or the updating matrix using Eq. \eqref{eq:w_from_phi}.

@@

For a non-periodic variable,
we have $\lambda_n = (-1)^{n-K}$
from Eqs. \eqref{eq:lambda_bandpass} and \eqref{eq:lambdan},
and $\mu_l = (-1)^{n-K+l}/(2 \, n) + \mathrm{sinr}(K, l, 2 \, n)$.
%\begin{equation}
%  \mu_l
%  =
%  (-1)^{n-K+l} / ( 2 \, n )
%    +
%  \mathrm{sinr}(K, l, 2n)
%  .
%  \notag
%  %\label{eq:mu_sinc_refl}
%\end{equation}
%
The updating matrix is given by
$w_{ij} = \mathrm{sinr}(K, i-j, 2 \, n) + \mathrm{sinr}(K, i+j-1, 2 \, n)$.
%
\note{$w_{ij}$
  is free from the oscillatory term, $(-1)^{n-K+1}$,
  that appeared in the window function, since
  its contribution to $\mu_{i-j}$ to the updating matrix
  is canceled by its contribution to either $\mu_{i+j-1}$
  or $\mu_{i+j-2n-1}$,
  whichever exists, of the opposite sign.}
%

@@

\note{The table of symbols is listed in Table \ref{tab:symbols}.
  \begin{table*}
  \footnotesize
  \centering
  \rowcolors{1}{white}{LightGray}
  \setlength{\tabcolsep}{4pt} % column separation
  \caption{\label{tab:symbols}
    Table of symbols.}
  \begin{tabular}{l | p{12cm} }
    Symbol          &   Description \\
    \hline
    $\mathbf{A}$    &   Transition matrix. \\
    $\alpha$        &   Updating magnitude. \\
    $b$             &   Cutoff of the window function of homogeneous updating schemes. \\
    $C_M$           &   Integral of the mass function, $M(\bar q)$.  \\
    $\delta$        &   Dirac's or discrete $\delta$-function. \\
    $c$             &   Constant. \\
    $\Err, \Err_R, \Err_A$          &   Error, residual error, asymptotic error. \\
    $\ln f$         &   Updating factor in the WL algorithm.  \\
    $\phi_{ki}$     &   Eigenvectors of the updating matrix. \\
    $g$             &   $1$ for a periodic variable, or for $2$ a non-periodic one. \\
    $G_i$           &   Integral of the autocorrelation function of histogram fluctuation at bin $i$. \\
    $\Gamma_k$      &   Integral of the autocorrelation function of mode $k$. \\
    $h_i(t)$        &   Instantaneous histogram.  \\
    $H_i$           &   Average normalized histogram.  \\
    $i, j$          &   Bin indices. \\
    $\lambda_k$     &   The $k$th eigenvalue of the updating matrix. \\
    $k, l$          &   Mode indices. \\
    $K$             &   Cutoff wave number of bandpass updating schemes.  \\
    $M(q)$          &   Mass function.   \\
    $m(q)$          &   Mass distribution,
                        $m(q) \equiv M(q)/\int_0^{ q(T) } M(q') \, dq'$.  \\
    $\mu_i$         &   Unadjusted window function of a homogeneous updating scheme. \\
    $n$             &   Number of bins. \\
    $\nu_k$         &   $\nu_k \equiv \lambda_k / \lambda$. \\
    $\rho_i$        &   Flat sampling distribution. \\
    $\pi_i$         &   (Instantaneous) distribution. \\
    $q(t)$          &   $\int_0^t \alpha(t') \, dt'$.  \\
    $\bar q(t)$     &   $q(T) - q(t)$.  \\
    $t, \tau$       &   Time. \\
    $T$             &   Simulation length. \\
    $\vartheta_k(q')$       &   $\vartheta_k(q') \equiv e^{\lambda_k \, [q - q(T)]}$. \\
    $u_i(t)$        &   Original bias potential. \\
    $v_i(t)$        &   Shifted bias potential. \\
    $\mathbf w$     &   Updating matrix. \\
    $v_{*i}(t)$     &   Bin-average-deducted bias potential, Eq. \eqref{eq:x_def}. \\
    $\xi_{*i}(t)$   &   Bin-average-deducted The noise of histogram. \\
    ${\tilde v}_k$  &   Mode of the shifted bias potential. \\
    ${\tilde u}_k$  &   Mode of the original bias potential. \\
    $z$             &   Reaction coordinate. \\
    $\xi_i(t)$    &   The noise part of the instantaneous histogram, $h_i(t)$.
  \end{tabular}
  \end{table*}
}
